Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2022-12-05 paper/abstract.tex
Revise the introduction as per reviewer comments

\newjunk{%
Avionics need to be engineered to be cyber-resilient meaning that systems are able to detect and recover from attacks or safely shutdown.
%
As there are few development tools for cyber-resiliency, designers rely on guidelines and checklists, sometimes missing vulnerabilities until late in the process where remediation is expensive.
%
Our solution is a model-based approach with cyber-resilience-improving transforms that insert high-assurance components such as filters to block malicious data or monitors to detect and alarm anomalous behavior.
%
Novel is our use of model checking and a verified compiler to specify, verify, and synthesize these components.
%
We define \emph{code contracts} as formal specifications that designers write for high-assurance components, and \emph{test contracts} as tests to validate their behavior.
%
A model checker proves whether or not code contracts satisfy test contracts in an iterative development cycle.
%
The same model checker also proves whether or not a system with the inserted components, assuming they adhere to their code contracts, provide the desired cyber-resiliency for the system.
%
We define an algorithm to synthesize implementations for code contracts in a semantics-preserving way that is backed by a verified compiler.
%
The entire workflow is implemented as part of the open source \emph{\brfcs} toolkit.
%
We report on our experience using \brfcs\ with a case study on a UAV system that is transformed to be cyber-resilient to communication and supply chain cyber attacks.
%
Our case study demonstrates that synthesizing correct implementations from code
contracts is feasible in real-world systems engineering.
}

>>>>

\newjunk{%
Avionics need to be engineered to be cyber-resilient meaning that systems are able to detect and recover from attacks or safely shutdown.
%
As there are few development tools for cyber-resiliency, designers rely on guidelines and checklists, sometimes missing vulnerabilities until late in the process where remediation is expensive.
%
Our solution is a model-based approach with cyber-resilience-improving transforms that insert high-assurance components such as filters to block malicious data or monitors to detect and alarm anomalous behavior.
%
Novel is our use of model checking and a verified compiler to specify, verify, and synthesize these components.
%
We define \emph{code contracts} as formal specifications that designers write for high-assurance components, and \emph{test contracts} as tests to validate their behavior.
%
A model checker proves whether or not code contracts satisfy test contracts in an iterative development cycle.
%
The same model checker also proves whether or not a system with the inserted components, assuming they adhere to their code contracts, provide the desired cyber-resiliency for the system.
%
We define an algorithm to synthesize implementations for code contracts in a semantics-preserving way that is backed by a verified compiler.
%
The entire workflow is implemented as part of the open source \emph{\brfcs} toolkit.
%
We report on our experience using \brfcs\ with a case study on a UAV system that is transformed to be cyber-resilient to communication and supply chain cyber attacks.
%
Our case study demonstrates that writing code contracts and then synthesizing correct implementations from them is feasible in real-world systems engineering for cyber-resilience.<
}

### 2022-12-05 paper/intro.tex
Revise the introduction as per reviewer comments

We have been developing the {\em BriefCASE} toolsuite to address this need.
\brfcs\ is a Model-Based Systems Engineering (MBSE) environment
built in the Open Source AADL\footnote{AADL is the acronym for
Architecture Analysis and Design Language~\cite{aadl}.}  Tool
Environment (OSATE) to add new capabilities for building
cyber-resilient systems.

In this paper we describe how \brfcs\ facilitates specifying, testing,
verifying, and synthesizing high assurance components inserted into a
system to improve its cyber-resiliency.  The main organizing concept
is that of an architecture-to-architecture \emph{security-improving
transform}, achieved via the insertion of a new architectural
component aimed at mitigating a cyber-vulnerability.  We describe two
such transformations in this paper: (1) the insertion of a
filter to prevent malformed data from a malicious actor being
propagated to downstream components, and (2) the insertion of a
runtime monitor to detect (and alert) unexpected behaviors arising
from untrusted components.

{\agr}
\footnote{AGREE is the acronym for \emph{Assume Guarantee Reasoning Environment} \cite{agree2013}.}
provides the formal setting for our work.
{\agr} is implemented in OSATE as an \emph{annex}; hence AGREE
specifications inherit AADL constructs for describing system
structure, interconnections, and data. {\agr} uses \emph{contracts}
expressed in a variant of Past-Time Linear Temporal Logic to formally
specify input and output properties of components and systems. In this
paper we introduce the notion of a \emph{code contract}: a class of
contracts that support the synthesis of implementations. The code
contract language is Turing complete, allowing the designer to specify
arbitrarily complex behavior.  These contracts are unit tested for
correctness with \emph{test contracts}.  Test contracts define test
scenarios to be implemented by the code contract under test.  {\agr}
proves whether or not that code contract implementation is correct,
enabling the designer to iteratively test the component behavior
inside the \brfcs\ environment.  Once the behavior of the code
contract is verified, \agr\ proves that---due to the newly included
high-assurance components---the hardened system meets its
cyber-resiliency requirements.

A novel aspect of our work is the synthesis of {\ckml} code from the
code contracts for high-assurance components. {\ckml} is a dialect of
the ML programming language possessing a verified
compiler \cite{cakeml}.  We give a detailed outline of the synthesis
path from code contracts to \ckml\ code, providing a formal framework
in which to argue correctness. \ckml\ then provides a verified
compilation path to several different target binaries (and also
proving that the meaning of the \ckml\ source code is exactly
preserved in the final binaries).  A key contribution here is that the
code contract semantics are defined in such a way that the \agr\
verification results for the code contract hold for the deployed
component, \ie, the component will detect and prevent the indicated
cyber-vulnerabilities over all possible finite inputs. Preliminary
work has shown how to extend our work to handle infinite input traces
as these systems are inherently reactive and intended to run
forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.

@@

As indicated in the preceding discussion, {\brfcs} is aimed at
addressing the lack of tool support for the problem of enforcing
and implementing the class of \emph{cyber-resilience} system
requirements.  Our approach to the problem is embedded in a
substantial collection of tools to be wholistic in addressing cyber-resilience at the system level, so it is important to clarify the
research impact. We see the following as the distinguishing
contributions of this paper:

\begin{compactitem}
  \item The language and semantics of \emph{code contracts} provide a
  means to join system-level contract analysis results (arising from
  model checking) with component implementation correctness results
  (arising from property-preserving code generation).

\item We provide a strong formal path connecting (A) top-level system cyber-resilience
  requirements, (B) specifications for high-assurance components
  inserted to mitigate against specific cyber-resilience problems, and
  (C) synthesis to a programming language with a verified compiler.

@@

@@

\begin{comment}
  BriefCASE incorporates model-level cyber analysis tools (presently
  GearCASE~\cite{gearcase2020} and DCRYPPS~\cite{dcrypps2019}) which can
  examine AADL models for potential vulnerabilities and suggest
  cyber-security requirements to mitigate them.  A library of
  architectural transforms guides the system engineer through automated
  model transformations that modify the architecture to address these
  requirements, possibly inserting new high-assurance components into
  the system.  Implementations for the new components are synthesized
  from formal specifications using
  SPLAT~\cite{slind-hcss2020},~\cite{formal-filter-synth-langsec}
  (Semantic Properties for Language and Automata Theory).

  Formal
  verification that the transformed system model meets its requirements
  is accomplished via \agr~\cite{agree2013} (Assume Guarantee Reasoning
  Environment).
  %The AGREE analysis \emph{assumes} properties on the inputs of a given component of the system, and attempts to formally prove the conjectured \emph{guarantees} of the output.
  \agr\ is a {\em compositional assume-guarantee} style model checker
  for AADL models that attempts to prove properties about one layer of
  an architecture using properties allocated to its subcomponents.
  Cyber-resilient code implementing the verified model is then
  automatically generated using the High Assurance Modeling and Rapid
  Engineering for Embedded Systems (HAMR) toolkit~\cite{hamr}.  If
  desired, this code can be targeted to the formally verified seL4
  secure microkernel~\cite{sel4-2009}.
\end{comment}

>>>>

We have been developing the {\em BriefCASE} \oldjunk{toolsuite} toolkit to address this need.
\brfcs\ is a Model-Based Systems Engineering (MBSE) environment
built in the Open Source AADL\footnote{AADL is the acronym for
Architecture Analysis and Design Language~\cite{aadl}.}  Tool
Environment (OSATE) to add new capabilities for building
cyber-resilient systems. 
%
In this paper we describe how it facilitates specifying, testing,
verifying, and synthesizing high assurance components inserted into a
system to improve its cyber-resiliency.  The main organizing concept
is that of an architecture-to-architecture \emph{security-improving
transform}, achieved via the insertion of \oldjunk{a} new architectural
component\newjunk{s} aimed at mitigating \oldjunk{a} cyber-vulnerabilit\oldjunk{y}\newjunk{ies}.  We describe two
such transformations in this paper: (1) the insertion of a
filter to prevent malformed data from a malicious actor being
propagated to downstream components, and (2) the insertion of a
runtime monitor to detect (and alert) unexpected behaviors arising
from untrusted components.

\oldjunk{%
{\agr}
provides the formal setting for our work.
{\agr} is implemented in OSATE as an \emph{annex}; hence AGREE
specifications inherit AADL constructs for describing system
structure, interconnections, and data. {\agr} uses \emph{contracts}
expressed in a variant of Past-Time Linear Temporal Logic to formally
specify input and output properties of components and systems. In this
paper we introduce the notion of a \emph{code contract}: a class of
contracts that support the synthesis of implementations. The code
contract language is Turing complete, allowing the designer to specify
arbitrarily complex behavior.  These contracts are unit tested for
correctness with \emph{test contracts}.  Test contracts define test
scenarios to be implemented by the code contract under test.  {\agr}
proves whether or not that code contract implementation is correct,
enabling the designer to iteratively test the component behavior
inside the \brfcs\ environment.  Once the behavior of the code
contract is verified, \agr\ proves that---due to the newly included
high-assurance components---the hardened system meets its
cyber-resiliency requirements.
}

\newjunk{%
The contributions of this paper relate to the aspects of \brfcs\ concerning evidence that a high-assurance component meets its intended purpose and is properly implemented. 
%
We leverage the existing {\agr}\footnote{AGREE is the acronym for \emph{Assume Guarantee Reasoning Environment}.} model checker to provide evidence of the former and the existing verified \ckml\ compiler to provide evidence of the latter \cite{agree2013,cakeml}. 
%
Specifically, designers write \emph{contracts} to formally specify input and output properties of the system and its components, and \agr\ proves via model checking whether or not the system composition preserves contract requirements at component inputs and guarantees contract obligations at system outputs.
%
In this way, \agr\ provides the formal framework we use to describe the behavior of an inserted high-assurance component, such as a filter or monitor, and then prove if adding such a component to the system improves
cyber-resiliency.
}

\newjunk{
We then use semantics preserving transformations to rewrite contracts into \ckml\ where the \ckml\ compiler then turns such contracts into binaries for the intended target platforms.
%
The verified compiler, with the semantics preserving transforms, ensures that the high-assurance component adheres to its contract specification.
%
This assurance fulfills the underlying assumption of the \agr\ model checker's proof: a component implementation adheres to its contract.
}

\newjunk{
As such, this paper defines a
\emph{code contract} language that is a subset of the \agr\ language
but with sequential evaluation semantics as opposed to \agr's existing
parallel semantics.  The sequential semantics are requisite to
code synthesis and are defined such that \agr\ is able to still prove the contract composition of the system works.
This paper further defines a notion of \emph{test contract} for unit
testing code contracts outside of the system composition to verify
input to output behavior as well as general properties of the code
contract.  
%
Intuitively, the behavior of a high assurance component is
defined by the designer in a code contract, the designer then writes test contracts to validate that behavior by asking \agr\ to prove if the code contract ensures the output obligations in the test contract under the test contracts input assumptions.
%
Once the designer is satisfied that the code
contract is correct, the system composition, with its various contracts, is verified to conclude
that the system meets its cyber-requirements.
}

\newjunk{%
Showing that a component is properly implemented is accomplished
through formal synthesis from the code contract.  This paper presents the Semantic Properties
of Language and Automata Theory (\splt) tool that generates
\ckml\ code to implement the code contract.  \ckml\ is a
functional language that includes additional proofs and tools built
around it \cite{cakeml}.  The \ckml\ compiler itself is verified to
generate binary implementations that are equivalent in semantics to
the original input program.  That proof ensures that the resulting
binary exactly preserves the meaning of the original code contract and
assures that the component is faithfully implemented over all possible finite inputs. 
Preliminary
work has shown how to extend our work to handle infinite input
since most systems of interest are inherently reactive and intended to run
forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.
}

\oldjunk{%
A novel aspect of our work is the synthesis of {\ckml} code from the
code contracts for high-assurance components. {\ckml} is a dialect of
the ML programming language possessing a verified
compiler.  We give a detailed outline of the synthesis
path from code contracts to \ckml\ code, providing a formal framework
in which to argue correctness. \ckml\ then provides a verified
compilation path to several different target binaries (and also
proving that the meaning of the \ckml\ source code is exactly
preserved in the final binaries).  A key contribution here is that the
code contract semantics are defined in such a way that the \agr\
verification results for the code contract hold for the deployed
component, \ie, the component will detect and prevent the indicated
cyber-vulnerabilities over all possible finite inputs.
}

@@

As indicated in the preceding discussion, {\brfcs} is aimed at
addressing the lack of tool support for the problem of enforcing
and implementing the class of cyber-resilience system
requirements.  Our approach to the problem is embedded in a
substantial collection of tools to be wholistic in addressing cyber-resilience at the system level, so it is important to clarify the
research impact. We see the following as the distinguishing
contributions of this paper:

\begin{compactitem}
  \item The language and semantics of \emph{code contracts} \oldjunk{provide} \newjunk{as} a
  means to join system-level contract analysis results (arising from
  model checking) with component implementation correctness results
  (arising from property-preserving code generation).

@@

  \item A strong formal path connecting (A) top-level system cyber-resilience
  requirements, (B) specifications for high-assurance components
  inserted to mitigate against specific cyber-resilience problems, and
  (C) synthesis to a programming language with a verified compiler.

@@
